---
title: "p8105_hw3_sz3214"
author: "Shizhe Zhang"
date: "2024-10-08"
output: html_document

---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# Problem 1

**load data**
```{r}
library(p8105.datasets)
data("ny_noaa")
```

**Overall**
```{r}
str(ny_noaa)
colMeans(is.na(ny_noaa)) * 100
```

The dataset consists of 2,595,176 rows and 7 columns. Each row represents a daily weather observation, and the key variables include:

`id`: A character variable identifying the weather station.
`date`: A date variable representing the date of the observation.
`prcp`: An integer variable indicating precipitation.
`snow`: An integer variable for snowfall measurements.
`snwd`: An integer variable for snow depth.
`tmax`: A character variable representing the maximum temperature for the day.
`tmin`: A character variable representing the minimum temperature.

Several key variables contain many NA values: 

`prcp`: 5.62% missing value.
`snow`: 14.69% missing value.
`snwd`: 22.80% missing value.
`tmax`: 43.71% missing value.
`tmin`: 43.71% missing value.

**data cleaning**
```{r}
ny_noaa = ny_noaa |> 
  janitor::clean_names() |> 
  mutate(year = format(date, "%Y"),
         month = format(date, "%m"),
         day = format(date, "%d"),
         tmax = as.integer(tmax) / 10,
         tmin = as.integer(tmin) / 10,
         prcp = prcp / 10)

ny_noaa |> 
  group_by(snow) |> 
  summarise(count = n()) |> 
  arrange(desc(count)) |> 
  head(n=5)
```
The most commonly observed value for snowfall is 0. 



